---
title: "Toronto COVID19 Cases Exploratory Data Analysis"
author: "Kwan Ho (Louis) Lee"
date: "2023-05-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction

This exercise aims to explore the COVID-19 Cases data across Toronto, the original data set was extracted from the Open Data Portal of City of Toronto: https://open.toronto.ca/dataset/covid-19-cases-in-toronto/. This Exploratory Data Analysis leverages the Tidyverse package for data manipulation and visualization.

### Install and Load Tidyverse Pacakge
```{r Install tidyverse}
install.packages("tidyverse",repos = "http://cran.us.r-project.org")
library("tidyverse")
```

### Set Working Directory
```{r Set Working Directory}
setwd("/Users/lou/Desktop/R Dataset")
getwd()
```

### Import Dataset From the Original CSV File
```{r Obtain Dataset}
data = read_csv("COVID_YYZ.csv")
glimpse(data)
str(data)
colnames(data)
head(data)
```

### Trend of COVID19 Cases in Toronto
```{r Plot Line Chart}
# Process data summary
daily_caseno <- data %>%
  group_by(Episode_Date) %>%
  summarise(Daily_new_cases = (count = n()))
# Convert Data Type of the Episode Date column
daily_caseno$Episode_Date <- as.Date(daily_caseno$Episode_Date, "%d/%m/%Y")
# Plot the line chart
ggplot(data = daily_caseno, mapping = aes(x = Episode_Date, y = Daily_new_cases)) +
  geom_line() +
  labs(x = "", y = "No. of Daily New Cases", title = "COVID19 Trend in Toronto", 
       caption = "Data Source: Toronto Public Health")
```

### Distribution of COVID19 Cases by Age
```{r Plot Pie Chart}
# Replace NA values
data$Age_Group[is.na(data$Age_Group)] <- "Not Specified"
# Process data summary
age_dist <- data %>%
  group_by(Age_Group) %>%
  summarise(Count = n()) %>%
  mutate(Percentage = round((Count / sum(Count) * 100), 2)) %>%
  mutate(ypos = cumsum(Percentage) - 0.5 * Percentage + 1)
# Set the fill colors of the pie chart
colorgroup <- c("pink", "red", "brown", "orange", "yellow2", "green", "skyblue", "blue", "purple", "grey")
# Plot the pie chart
ggplot(data = age_dist, mapping = aes(x = "", y = Count, fill = Age_Group)) +
  geom_bar(color = "white", stat = "identity", width = 1) +
  coord_polar("y", start = 180) +
  theme_void() +
  scale_fill_manual(values = colorgroup) +
  geom_text(aes(label = paste0(Percentage, "%")), size = 5, color = "white", 
            position = position_stack(vjust = 0.6)) +
  labs(title = "Age Distribution for COVID19 Patients in Toronto", 
       caption = "Data Source: Toronto Public Health")
```

### Distribution of COVID19 Cases by Location
```{r Plot Bar Chart}
# Process data summary
location_dist <- data %>%
  group_by(Neighbourhood_Name) %>%
  summarise(Count = n()) %>%
  arrange(-Count)
# Remove NA rows
location_dist <- location_dist[-1,]
location_dist <- head(location_dist, 10)
# Plot the bar chart
ggplot(data = location_dist, mapping = aes(x = reorder(Neighbourhood_Name, -Count), y = Count)) +
  geom_col(fill = "red") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(x = "Neighbourhood", y = "No. of COVID19 Cases", 
       title = "Distribution of COVID-19 Cases Across Toronto", caption = "Data Source: Toronto Public Health")
```

### Distribution of COVID19 Cases by Age and Gender
```{r Plot Heat Map}
# Group gender values
data$Client_Gender[data$Client_Gender == "NOT LISTED, PLEASE SPECIFY" |
                     data$Client_Gender == "OTHER" |
                     data$Client_Gender == "TRANS MAN" |
                     data$Client_Gender == "TRANS WOMAN" |
                     data$Client_Gender == "TRANSGENDER"] <- "NON-BINARY"
# Process data summary
group <- data %>%
  group_by(Age_Group, Client_Gender) %>%
  summarise(Count = n(), .groups = "keep")
# Plot the heat map
ggplot(data = group, mapping = aes(x = Age_Group, y = Client_Gender, fill = Count)) +
  geom_tile() + 
  scale_fill_gradient(low = "yellow", high = "red") +
  theme_classic() +
  labs(x = "Age Group", y ="No. of Patient", title = "COVID19 Patient Population by Gender and Age", 
       caption = "Data Source: Toronto Public Health" ) +
  theme(axis.text.x = element_text(angle = 45, hjust=1))
```

  
